---
title: "A week without digital notes. The after"
description: "A micro-update nobody asked for."
date: 2026-06-18
categories: [note-taking, PKM, microblogging]
draft: true
---

A few months ago I spent a week without digital notes and measured my notebook like it owed me something. The post is [here](https://cecibaldoni.github.io/blog/week-without-digital.html), if you missed it.

I said at the end that I changed almost nothing. That was true. But "almost" has been sitting in the back of my head.

So here's the data.

## Bullet points, the vengeance

```{r}
#| echo: false
#| message: false

library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)

notebook_data <- readRDS("../data/notebook_data.rds")

col_before <- "#7a8c78"
col_during <- "#c46e3a"
col_after  <- "#4a7a8a"
palette    <- c("Before" = col_before, "During" = col_during, "After" = col_after)

base_theme <- theme_void(base_size = 15) +
  theme(plot.title       = element_text(face = "bold", size = 16, margin = margin(b = 10)),
        axis.text.x      = element_text(size = 10, color = "grey30", margin = margin(t = 8)),
        axis.text.y      = element_text(size = 10, color = "grey30", margin = margin(r = 8)),
        axis.title.y     = element_text(size = 12, color = "grey40", angle = 90, margin = margin(r = 10)),
        legend.position  = "none",
        panel.grid.major = element_line(color = "grey92"),
        plot.background  = element_rect(fill = "#fdfbf7", color = NA),
        panel.background = element_rect(fill = "#fdfbf7", color = NA))

bullets_summary <- notebook_data %>%
  group_by(period) %>%
  summarise(mean_bullets = mean(bullets), .groups = "drop")

ggplot(bullets_summary, aes(x = period, y = mean_bullets, fill = period)) +
  geom_col(width = 0.5, alpha = 0.9) +
  geom_text(aes(label = round(mean_bullets, 1)),
            vjust = -0.4, size = 3.5, color = "grey30") +
  scale_fill_manual(values = palette) +
  labs(title = "Bullet points per day",
       y = "Mean", x = NULL) +
  base_theme +
  plot_annotation(caption = "The notebook learned to make lists")
```

Bullet point usage went way up. This was not a conscious decision, and I have no tidy explanation for this.

The only thing that I associate this with is the Logseq structure. Not only I started using way more bullet points to resamble the block logic in Logseq, but I even started making *nested* bullet points. 

But then why I did not use any during my digital detox week, and more when Logseq was available? These kind questions keep me up at night.


## But what about pages used?

This data didn't make it into the original post, but someone asked me about this so you're getting it now.


```{r}
#| echo: false
#| message: false
#| warning: false

bridge_1 <- data.frame(day = 7.5,  period = factor("Before", levels = c("Before", "During", "After")), pages = NA)
bridge_2 <- data.frame(day = 14.5, period = factor("During", levels = c("Before", "During", "After")), pages = NA)

pages_line <- notebook_data %>%
  select(day, period, pages) %>%
  bind_rows(bridge_1, bridge_2) %>%
  arrange(day)

ggplot(pages_line, aes(x = day, y = pages, group = 1)) +
  annotate("rect", xmin = 7.5,  xmax = 14.5, ymin = -Inf, ymax = Inf,
           fill = "#f5ede4", alpha = 0.5) +
  annotate("rect", xmin = 14.5, xmax = 21.5, ymin = -Inf, ymax = Inf,
           fill = "#e4edf5", alpha = 0.3) +
  geom_line(color = "grey80", linewidth = 2) +
  geom_line(aes(color = period), linewidth = 2) +
  geom_point(aes(color = period), size = 4) +
  scale_color_manual(values = palette) +
  scale_x_continuous(breaks = c(1:7, 8:14, 15:21),
                     labels = c(paste0("D", 1:7), paste0("D", 1:7), paste0("D", 1:7))) +
  labs(title = "Pages used per day", x = NULL, y = "Pages") +
  base_theme +
  plot_annotation(caption = "Questionable sample size.")

```

Analog usage went up. Not to the heights of the challenge week (that was a forced experiment, with no other outlet), but meaningfully above where it started. 

The notebook is getting more use than before. 

Whether that means the experiment worked or I just used the excuse to buy a new notebook, it will remain a mistery.


## In the end

I ended the [original post](https://cecibaldoni.github.io/blog/week-without-digital.html) saying that I changed almost nothing, because that is how I felt at the time. Immediately after the self-imposed return to technology, my life felt going back to its original axis.

More than 8 years in research should have taught me not to wave those statement without proof: the data says something changedso it must mean that something recalibrated. Which is either the best possible outcome of an experiment or a sign I should run more of them. 